---
title: "Database Workshop Workbook"
output: html_document
---

```{r, setup}
library(DBI)
library(dplyr)
library(ggplot2)

# Which function from DBI? 
  # Hint 1: Find the connection in the connections pane
  # Hint 2: Tab autocomplete shows you all functions in a package
con <- DBI::___(odbc::odbc(), "Postgres Prod")
```

# Examining Orders

How many orders are there (note: 1 row/order in the orders table)?

```{r}
ord <- tbl(con, dbplyr::in_schema("retail", "orders"))
# dplyr probably has a function for this...

```

Who are the 5 highest-ordering customers by volume? 

```{r}

# Might be worth examining data types, sometimes they're weird
# Perhaps as.numeric() would be helpful here
```

Can I make a histogram of their orders?
```{r}
# Some ggplot code that should work with the right input
ggplot(plot_dat, aes(x = customer_name, y = n)) +
  geom_col()
```

What's the longest name in the customer table?
```{r}

# dplyr::summarize might be useful here
# Also, perhaps the postgres string functions: https://www.tutorialspoint.com/postgresql/postgresql_string_functions.htm
# Some common functions have been translated...can't hurt to try!
```

Let's write a function to return the number of customers with an arbitrary number of 
letters in their name.

It's all good to go, but you need to fill in what goes in the filter.
```{r}
# Checkout ?nchar...and remember !! allows you to evaluate before sending

#' @param x a character
#' @return a number
#' @examples
cust_n_letters("hello world")
cust_n_letters <- function(x) {
  tbl(con, dbplyr::in_schema("retail", "customer")) %>%
    dplyr::filter(...) %>%
    dplyr::tally()
}
```

