---
title: "Database Workshop Workbook"
output: html_document
---

```{r, setup}
library(DBI)
library(dplyr)
library(ggplot2)

# Which function from DBI? 
  # Hint 1: Find the connection in the connections pane
  # Hint 2: Tab autocomplete shows you all functions in a package
con <- DBI::dbConnect(odbc::odbc(), "Postgres Prod")
```

# Examining Orders

How many orders are there (note: 1 row/order in the orders table)?

```{r}
ord <- tbl(con, dbplyr::in_schema("retail", "orders"))
# dplyr probably has a function for *count*ing the rows in a data frame...
# You can browse functions in a package by typing e.g. dplyr:: and then using arrow keys

dplyr::count(ord)
```

What are the names of the 5 highest-ordering customers by volume? 
Challenge Q: Names of the 5 highest-ordering customers by revenue?

```{r}
# The orders table has one row per order, so each time a customer is in there is one order.
# You can ignore the step_id variable.

# Maybe customer names are in another table?
cust <- tbl(con, dbplyr::in_schema("retail", "customer"))

(plot_dat <- ord %>%
  count(customer_id) %>%
  inner_join(cust) %>%
  arrange(desc(n)))

# Challenge
line <- tbl(con, dbplyr::in_schema("retail", "line_item"))
prod <- tbl(con, dbplyr::in_schema("retail", "product"))
ord %>%
  inner_join(cust) %>%
  inner_join(line) %>%
  inner_join(prod) %>%
  group_by(customer_id) %>%
  summarize(total_rev = sum(price)) %>%
  arrange(desc(total_rev))
```

Can I make a histogram of their orders?
```{r}
plot_dat <- plot_dat %>%
  collect() %>%
  mutate(n = as.numeric(n))

# Some ggplot code that should work with the right input, maybe just print the data frame if it's not?
# Perhaps as.numeric() would be helpful here
ggplot(plot_dat, aes(x = customer_name, y = n)) +
  geom_col()
```

What's the longest name in the customer table?
```{r}
# dplyr::summarize might be useful here
# Also, perhaps the postgres string functions: https://www.tutorialspoint.com/postgresql/postgresql_string_functions.htm
# Some common functions (min, mean, max, sum) have been translated...

cust %>%
  mutate(l = str_length(customer_name)) %>%
  filter(l == max(l, na.rm = TRUE))
```

Let's write a function to return the customers with an arbitrary number of 
letters in their name.

It's all good to go, but you need to fill in what goes in the filter.
```{r}
# Checkout ?nchar...and remember !! allows you to evaluate before sending

#' @param x a character
#' @return a vector of names
#' @examples
cust_n_letters("hello world")
cust_n_letters <- function(x) {
  tbl(con, dbplyr::in_schema("retail", "customer")) %>%
    dplyr::filter(str_length(customer_name) == !!nchar(x)) %>%
    dplyr::pull(customer_name)
}
```

