---
title: "Exercise 2.2"
author: "Hadrien@rstudio.com"
date: "2/18/2020"
output: html_document
---


# Exercise 2.2: make an API call with `httr::GET()`

- Make a request by calling `httr::GET()` with the URL https://datausa.io/api/data?drilldowns=State&measures=Population&year=latest

```{r}
library(httr)

# save url
url <- "https://datausa.io/api/data?drilldowns=State&measures=Population&year=latest"

# make request and save response
resp <- ____(url)

```


- pull out important parts of the response with helper functions like `status_code()` and `headers()`
```{r}
# check response status code
_________(resp) 

# print all response headers
______(resp)

# check response type
_______(resp)

```


- retrieve the body of the request using `content()`

```{r}
______(resp)
```


```{r}
# convert character vector to R dataframe
library(jsonlite)
________(content(resp, as = "text"))
```


???
"HTTP methods like GET, POST, UPDATE, and DELETE inform the server about the action to be performed. Different HTTP methods invoked on the same URL provide different functionality..."

"Use GET for reading resources.  GET requests never, ever change the state of the resource. They have no side effect; the GET method has a read-only semantic. GET is idempotent. Consequently, you can cache the calls perfectly."

"httpbin.org ... accepts many types of http request and returns json that describes the data that it received. This makes it easy to see what httr is doing." - httr vignette

"GET() is used by your browser when requesting a page, and POST() is (usually) used when submitting a form to a server. " - httr vignette